Evaluating infectious disease forecasts \\ with allocation scoring rules

### 2024-02-17 alloscore_manuscript/alloscore-application-shorter.Rnw
convert supplement to appendices

\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

\usepackage{amsmath, amsfonts, amssymb}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{cases}
\usepackage{caption}
\usepackage{xr-hyper}
\usepackage{hyperref}
\usepackage{xurl}
\externaldocument{supplement_short}

@@

We begin with an informal description of the allocation score and some examples illustrating its key characteristics in
section \ref{sec:methods.overview}. In section \ref{sec:methods.detailed} we develop the allocation score more
carefully, using a decision theoretic procedure for deriving proper scoring rules. (See section \ref{sec:proper} of the
supplement for a definition of a proper scoring rule and a more technical discussion of the procedure). In section
\ref{sec:methods.related}, we note that another group of common scores including the quantile score, WIS, and CRPS, can
also be derived from decision theoretic foundations \textemdash starting from a different decision making context.

@@

\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$
and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$. The means of these distributions are
given by the scale parameters $\sigma_i$. When the marginal forecasts are exponential distributions, it can be shown
that the optimal allocation divides the available resources among the locations proportionally to the scale parameters
$\sigma_i$ (see section \ref{sec:bayes-quantiles} of the supplemental materials). If $K = 5$ units of the resource are
available, the optimal allocation according to $F$ would be 1 and 4 resource units in locations 1 and 2, respectively. 
Increasing $K$ to $10$ changes this optimal allocation to 2 and 8. Figure~\ref{fig:exp_alloc_example} illustrates the 
situation.

@@

We arrive through this discussion at a three-step procedure for developing scoring rules for probabilistic forecasts:
\begin{enumerate}
  \item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$
    eventually occurs.
  \item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under
    the distribution $F$.
  \item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: 
    $S(F, y) = s(x^F, y)$.
\end{enumerate}
This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss
function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} --- a terminology we find to
be convenient but which is not, to our knowledge, standard.  In section \ref{sec:proper} of the supplement we review the
basic fact from the scoring rule literature (e.g., \cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring
rules are proper by construction.

@@

In the supplement we show that the components of the Bayes act are quantiles $x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a
probability level $\tau^{F,K}$ that depends on the forecast $F$ and the resource constraint $K$, but is shared across
all locations. This probability level is the level at which the resource constraint is satisfied: $\sum_{i=1}^N
F_i^{-1}(\tau^{F,K}) = K$. This tells us that in order to allocate optimally (according to $F$), we must divide
resources among the locations so that there is an equal forecasted probability in every location that the allocation is
sufficient to meet resource need. This solution to the allocation problem is well-known in inventory management and is
often attributed to \cite{hadleywhitin1963}.

@@

Submission to the Forecast Hub followed a weekly cycle, and each Monday the Hub collected the most recent forecasts
submitted by all teams that met certain inclusion criteria and created ensemble forecasts using quantile averaging
\citep{ray_comparing_2023}. Our analysis includes these ensembles (COVIDhub-ensemble and COVIDhub-trained\_ensemble) as
well as one other ensemble of hub models created by another team (JHUAPL-SLPHospEns) and several other individual
models. Models were eligible to be included in the analysis if they were designated as a ``primary'' model from a team.
For a model to have a complete, eligible submission in a given week, it had to have a 14 day-ahead forecast for all 50
states plus Washington DC. Models had to have a complete forecast for at least 4 of the 15 weeks in the analysis to be
eligible for inclusion.

@@

As described above, predictions were submitted to the Forecast Hub in the form of a set of 23 quantiles of the
predictive distribution. The WIS can be directly calculated from these quantiles. However, calculation of the AS
requires that the full cumulative distribution functions of the forecast distributions for each location are available.
For the purpose of this analysis, to calculate the allocation score we approximated the full cumulative distribution
functions based on the provided quantiles. On the interior of the provided quantiles we used a monotonic cubic spline to
interpolate the quantiles, and in the lower and upper tails we used normal distributions with parameters selected so as
to match the two lowest and two highest quantiles (see section \ref{sec:distfromq} the supplement for further details).
In section \ref{sec:propriety_of_parametric_approximation} of the supplement, we show that if this evaluation procedure
had been specified prospectively, the resulting score would be proper, \textemdash but that a post hoc application of this
procedure is improper. We use the procedure here to illustrate the properties of the AS, and note that a collaborative
forecasting hub interested in using the AS for evaluation could circumvent propriety issues by communicating the
definition of the AS to forecasters and then collecting allocations at specified resource levels as part of forecast
submissions.

@@

Of the 10 models evaluated for this one week, the CU-select model had the most accurate forecasts according to the
allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table
\ref{tab:multi-k-scores}). The JHUAPL-Bucky model had the second best MWIS but the third worst allocation score.

@@

A state-wise comparison using Figures \ref{fig:thermometer-plot} and \ref{fig:multi-loc-ranks}A of the JHUAPL-Bucky and
CU-select models shows that while the JHUAPL-Bucky forecast distributions were more centered on the eventual
observations in many states, the suggested allocations of the JHUAPL-Bucky model were less efficient than those of the
CU-select model. Thus JHUAPL-Bucky achieved a worse allocation score than CU-select (on this particular week) by
allocating excess resources to several states, such as Pennsylvania and Michigan, rather than to states such as Florida
and California where these resources would have reduced unmet need.  These allocation errors of JHUAPL-Bucky resulted
from forecasts failing to consistently capture the relative resource needs across different states. The CU-select model
made some similar errors — most prominently, over-allocating resources to Ohio — but overall, it more successfully
forecasted the resource demands across different locations in relative terms.

On the other hand, CU-select had worse performance as measured by MWIS. Its forecasts were biased downwards, and it
consistently incurred a large penalty for underprediction (Figure \ref{fig:multi-loc-ranks}B). The JHUAPL-Bucky model
had wider predictive intervals which more often included the observed level of daily hospital admissions. Its MWIS was
therefore less severely penalized by under- or over-predictions of the actual hospitalization counts.

<<multi-loc-ranks, fig.width=8, fig.cap="Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (JHUAPL-Bucky and CU-select). Panel A shows the observed resource need, in this case the observed number of hospitalizations, for each state, along with the hypothetical number of resources allocated to the given location based on the forecasts from each model. The number of available resources was fixed at 15,000 and forecasts from each model were used to determine an optimal allocation strategy before the resource need was known. For most locations the resource need exceeded the resources allocated, indicated by the `observed resource need' bar being larger than the `allocation' bar. Panel B shows the breakdown of the weighted interval score (WIS) into components of underprediction, overprediction and dispersion. Larger values of WIS indicate more error, and the full WIS score for each location can be decomposed into the three components shown here.">>=
## date points to target_end_date of 2022-01-10 to highlight
ref_date_for_plot <- "2021-12-20"
models_to_plot <- c("JHUAPL-Bucky", "CU-select", "per-capita")

@@

<<metrics-over-time, fig.height=8, fig.width=8, fig.cap="Hospital admissions and evaluation metrics over time. Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022. These are the dates for which forecasts were made and evaluated. A horizontal dashed line at 15,000 shows the hypothetical resource constraint $K$. Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states. The $x$-axis corresponds to the date of the observation that a model's prediction was targeting (e.g., the date the forecast was made plus the forecast horizon). AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down). In Panel B, the per-capita allocation is drawn in a heavier solid line. Panel C shows the MWIS metric across weeks, averaged across all states. MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.">>=
p2 <- ggplot(alloscore_data_summaries,
             aes(x=target_end_date, y=alloscore, color=model, shape=model, linewidth = model)) +
  geom_point() + geom_line()+
  scale_x_date(limits=range(score_data_summaries$target_end_date)) +
  scale_color_manual(values = model_palette) +
  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4, 6)) +
  scale_linewidth_manual(values = model_linewidths) +
  ylab("Allocation score") +
  theme(legend.pos="none") +
  xlab(NULL)

@@

Averaging across all weeks, the ensemble forecast achieved a better mean allocation score (MAS) than two benchmark
methods. The COVIDhub-ensemble had the best MAS across all weeks, the per-capita allocation had the second best MAS, and
the COVIDhub-baseline model had the sixth best allocation (Table \ref{tab:multi-week-performance-summary}). No
individual model had a better MAS than the per-capita allocation.

Model rankings on mean allocation score (MAS) and time-averaged mean weighted interval score
(taMWIS) were broadly similar, with some places of disagreement. The four most accurate and the five least accurate models were the same
according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
(This comparison excludes the per-capita allocation which does not use forecasts and therefore cannot be assigned a MWIS.)
Notably, the JHUAPL-SLPHospEns model ranked best on taMWIS but had only the fourth best MAS.

<<table-score-summary-all-weeks, results='asis'>>=
all_scores |>
  filter( !(reference_date.x %in% c("2021-11-22", "2022-02-28"))) |>
  group_by(model) |>
  summarize(N = sum(!is.na(alloscore)),
            MAS = mean(alloscore, na.rm=TRUE),
            taMWIS = mean(mwis, na.rm=TRUE)) |>
  filter(N > 12) |>
  mutate(`taMWIS rank` = rank(taMWIS, na.last = "keep")) |>
  arrange(MAS) |>
  select(model, MAS, taMWIS, `taMWIS rank`) |>
  xtable::xtable(caption = "Time averaged mean weighted interval scores (taMWIS) and mean allocation scores (MAS) by model across 13 weeks. These results show the average performance across time for the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, as well
  as for a per-capita allocation. MWIS is not defined for the per-capita allocation which does not use forecasts.",
                 digits = 0,
                 label = "tab:multi-week-performance-summary") |>
  print(include.rownames = FALSE, NA.string = "-")
@

@@

We can gain some insight into the discordance between MAS and taMWIS by examining the correlation between time-specific AS and MWIS values of the various models in Figure
\ref{fig:metrics-correlation}. We find, for example, no clear association between the consistent
and high MWIS rank of the JHUAPL-SLPHospEns model and its highly variable AS rank. This contrasts 
with a clearly positive correlation between MWIS and AS ranks in other models such as 
Karlen-pypm and USC-SI\_kJalpha.
We also can see in more detail the consistently high performance COVIDhub-ensemble on both scores.
And while it did not submit forecasts for enough weeks to be included in Table \ref{tab:multi-week-performance-summary}, CMU-TimeSeries is an interesting example of a model that 
performed consistently well on AS but had only middling MWIS ranks.

@@

AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of
hospitalizations on January 3, 2021 as well as for the per-capita allocation (Figure \ref{fig:multi-k}A). These
calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of
\Sexpr{format(tot_y, big.mark = ",")} that day. Model ranks by AS were fairly stable across all $K$, and there appears
to be an substantial interval around the observed value over which ranks are constant.

@@

In practice, epidemiological forecasts are often directed at a diverse set of end-users. It may be easy for some of
these users to summarize the consequences of how they use a forecast with a numerical loss function, but for others,
such as officials deciding how best to update public ``situational awareness'' of an outbreak, this may be essentially
impossible. Even those forecast uses that are easily framed as expected loss minimization may differ enough that no
single score would be appropriate for all users. Ideally, targeted forecasting tools could be developed through close
collaboration between modelers and public health officials. However, this may only be possible in settings with
sufficient staffing on both an analytics and a public health team. Increasingly, collaborative modeling hubs are being
used to generate ``one-size-fits-all'' forecasts for many locations at once. But it could still be beneficial in these
settings to expand the set of scoring methods used by the hub to include scores, such as the AS, which are based on
specific public health decision problems.  This could help end users better understand the value of available forecasts
as inputs to their particular decision making contexts. (We note that issues of propriety are raised when forecasts are
evaluated with scoring rules for which they were not elicited or via parameters imputed by a hub after submission, both
of which are done in our application example.  We consider these issues briefly in supplement sections
\ref{sec:distfromq} and \ref{sec:propriety_of_parametric_approximation}.)

@@

>>>>

\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
\usepackage{appendix}
\usepackage{amsmath, amsfonts, amssymb}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{cases}
\usepackage{caption}
\usepackage{xr-hyper}
\usepackage{hyperref}
\usepackage{xurl}

@@

We begin with an informal description of the allocation score and some examples illustrating its key characteristics in
section \ref{sec:methods.overview}. In section \ref{sec:methods.detailed} we develop the allocation score more
carefully, using a decision theoretic procedure for deriving proper scoring rules. (See appendix \ref{sec:proper}
for a definition of a proper scoring rule and a more technical discussion of the procedure). In section
\ref{sec:methods.related}, we note that another group of common scores including the quantile score, WIS, and CRPS, can
also be derived from decision theoretic foundations \textemdash starting from a different decision making context.

@@

\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$
and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$. The means of these distributions are
given by the scale parameters $\sigma_i$. When the marginal forecasts are exponential distributions, it can be shown
that the optimal allocation divides the available resources among the locations proportionally to the scale parameters
$\sigma_i$ (see appendices \ref{sec:bayes-quantiles}). If $K = 5$ units of the resource are
available, the optimal allocation according to $F$ would be 1 and 4 resource units in locations 1 and 2, respectively. 
Increasing $K$ to $10$ changes this optimal allocation to 2 and 8. Figure~\ref{fig:exp_alloc_example} illustrates the 
situation.

@@

We arrive through this discussion at a three-step procedure for developing scoring rules for probabilistic forecasts:
\begin{enumerate}
  \item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$
    eventually occurs.
  \item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under
    the distribution $F$.
  \item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: 
    $S(F, y) = s(x^F, y)$.
\end{enumerate}
This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss
function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} --- a terminology we find to
be convenient but which is not, to our knowledge, standard.  In appendix \ref{sec:proper} we review the
basic fact from the scoring rule literature (e.g., \cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring
rules are proper by construction.

@@

In appendix \ref{sec:bayes-quantiles} we show that the components of the Bayes act are quantiles $x_i^{F,K} = F_i^{-1}(\tau^{F,K})$ at a
probability level $\tau^{F,K}$ that depends on the forecast $F$ and the resource constraint $K$, but is shared across
all locations. This probability level is the level at which the resource constraint is satisfied: $\sum_{i=1}^N
F_i^{-1}(\tau^{F,K}) = K$. This tells us that in order to allocate optimally (according to $F$), we must divide
resources among the locations so that there is an equal forecasted probability in every location that the allocation is
sufficient to meet resource need. This solution to the allocation problem is well-known in inventory management and is
often attributed to \cite{hadleywhitin1963}.

@@

Submission to the Forecast Hub followed a weekly cycle, and each Monday the Hub collected the most recent forecasts
submitted by all teams that met certain inclusion criteria and created ensemble forecasts using quantile averaging
\citep{ray_comparing_2023}. Our analysis includes these ensembles (\texttt{COVIDhub-ensemble} and \texttt{COVIDhub-trained\_ensemble}) as
well as one other ensemble of hub models created by another team (\texttt{JHUAPL-SLPHospEns}) and several other individual
models. Models were eligible to be included in the analysis if they were designated as a ``primary'' model from a team.
For a model to have a complete, eligible submission in a given week, it had to have a 14 day-ahead forecast for all 50
states plus Washington DC. Models had to have a complete forecast for at least 4 of the 15 weeks in the analysis to be
eligible for inclusion.

@@

As described above, predictions were submitted to the Forecast Hub in the form of a set of 23 quantiles of the
predictive distribution. The WIS can be directly calculated from these quantiles. However, calculation of the AS
requires that the full cumulative distribution functions of the forecast distributions for each location are available.
For the purpose of this analysis, to calculate the allocation score we approximated the full cumulative distribution
functions based on the provided quantiles. On the interior of the provided quantiles we used a monotonic cubic spline to
interpolate the quantiles, and in the lower and upper tails we used normal distributions with parameters selected so as
to match the two lowest and two highest quantiles (see appendix \ref{sec:distfromq} for further details).
In appendix \ref{sec:propriety_of_parametric_approximation}, we show that if this evaluation procedure
had been specified prospectively, the resulting score would be proper, \textemdash but that a post hoc application of this
procedure is improper. We use the procedure here to illustrate the properties of the AS, and note that a collaborative
forecasting hub interested in using the AS for evaluation could circumvent propriety issues by communicating the
definition of the AS to forecasters and then collecting allocations at specified resource levels as part of forecast
submissions.

@@

Of the 10 models evaluated for this one week, the CU-select model had the most accurate forecasts according to the
allocation score while the USC-SI\_kJalpha model had the most accurate forecasts based on MWIS (Table
\ref{tab:multi-k-scores}). The \texttt{JHUAPL-Bucky} model had the second best MWIS but the third worst allocation score.

@@

A state-wise comparison using Figures \ref{fig:thermometer-plot} and \ref{fig:multi-loc-ranks}A of the \texttt{JHUAPL-Bucky} and
\texttt{CU-select} models shows that while the \texttt{JHUAPL-Bucky} forecast distributions were more centered on the eventual
observations in many states, the suggested allocations of the JHUAPL-Bucky model were less efficient than those of the
\texttt{CU-select} model. Thus \texttt{JHUAPL-Bucky} achieved a worse allocation score than CU-select (on this particular week) by
allocating excess resources to several states, such as Pennsylvania and Michigan, rather than to states such as Florida
and California where these resources would have reduced unmet need.  These allocation errors of \texttt{JHUAPL-Bucky} resulted
from forecasts failing to consistently capture the relative resource needs across different states. The \texttt{CU-select} model
made some similar errors — most prominently, over-allocating resources to Ohio — but overall, it more successfully
forecasted the resource demands across different locations in relative terms.

On the other hand, \texttt{CU-select} had worse performance as measured by MWIS. Its forecasts were biased downwards, and it
consistently incurred a large penalty for underprediction (Figure \ref{fig:multi-loc-ranks}B). The \texttt{JHUAPL-Bucky} model
had wider predictive intervals which more often included the observed level of daily hospital admissions. Its MWIS was
therefore less severely penalized by under- or over-predictions of the actual hospitalization counts.

<<multi-loc-ranks, fig.width=8, fig.cap="Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (\\texttt{JHUAPL-Bucky} and \\texttt{CU-select}). Panel A shows the observed resource need, in this case the observed number of hospitalizations, for each state, along with the hypothetical number of resources allocated to the given location based on the forecasts from each model. The number of available resources was fixed at 15,000 and forecasts from each model were used to determine an optimal allocation strategy before the resource need was known. For most locations the resource need exceeded the resources allocated, indicated by the `observed resource need' bar being larger than the `allocation' bar. Panel B shows the breakdown of the weighted interval score (WIS) into components of underprediction, overprediction and dispersion. Larger values of WIS indicate more error, and the full WIS score for each location can be decomposed into the three components shown here.">>=
## date points to target_end_date of 2022-01-10 to highlight
ref_date_for_plot <- "2021-12-20"
models_to_plot <- c("JHUAPL-Bucky", "CU-select", "per-capita")

@@

<<metrics-over-time, fig.height=8, fig.width=8, fig.cap="Hospital admissions and evaluation metrics over time. Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022. These are the dates for which forecasts were made and evaluated. A horizontal dashed line at 15,000 shows the hypothetical resource constraint $K$. Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states. The $x$-axis corresponds to the date of the observation that a model's prediction was targeting (e.g., the date the forecast was made plus the forecast horizon). AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down). In Panel B, the \\texttt{per-capita} allocation is drawn in a heavier solid line. Panel C shows the MWIS metric across weeks, averaged across all states. MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.">>=
p2 <- ggplot(alloscore_data_summaries,
             aes(x=target_end_date, y=alloscore, color=model, shape=model, linewidth = model)) +
  geom_point() + geom_line()+
  scale_x_date(limits=range(score_data_summaries$target_end_date)) +
  scale_color_manual(values = model_palette) +
  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4, 6)) +
  scale_linewidth_manual(values = model_linewidths) +
  ylab("Allocation score") +
  theme(legend.pos="none") +
  xlab(NULL)

@@

Averaging across all weeks, the ensemble forecast achieved a better mean allocation score (MAS) than two benchmark
methods. The \texttt{COVIDhub-ensemble} had the best MAS across all weeks, the \texttt{per-capita} allocation had the second best MAS, and
the \texttt{COVIDhub-baseline} model had the sixth best allocation (Table \ref{tab:multi-week-performance-summary}). No
individual model had a better MAS than the per-capita allocation.

Model rankings on mean allocation score (MAS) and time-averaged mean weighted interval score
(taMWIS) were broadly similar, with some places of disagreement. The four most accurate and the five least accurate models were the same
according to both metrics, although not in exactly the same order (Table \ref{tab:multi-week-performance-summary}).
(This comparison excludes the \texttt{per-capita} allocation which does not use forecasts and therefore cannot be assigned a MWIS.)
Notably, the \texttt{JHUAPL-SLPHospEns} model ranked best on taMWIS but had only the fourth best MAS.

<<table-score-summary-all-weeks, results='asis'>>=
all_scores |>
  filter( !(reference_date.x %in% c("2021-11-22", "2022-02-28"))) |>
  group_by(model) |>
  summarize(N = sum(!is.na(alloscore)),
            MAS = mean(alloscore, na.rm=TRUE),
            taMWIS = mean(mwis, na.rm=TRUE)) |>
  filter(N > 12) |>
  mutate(`taMWIS rank` = rank(taMWIS, na.last = "keep")) |>
  arrange(MAS) |>
  select(model, MAS, taMWIS, `taMWIS rank`) |>
  xtable::xtable(caption = "Time averaged mean weighted interval scores (taMWIS) and mean allocation scores (MAS) by model across 13 weeks. These results show the average performance across time for the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, as well
  as for a per-capita allocation. MWIS is not defined for the \\texttt{per-capita} allocation which does not use forecasts.",
                 digits = 0,
                 label = "tab:multi-week-performance-summary") |>
  print(include.rownames = FALSE, NA.string = "-")
@

@@

We can gain some insight into the discordance between MAS and taMWIS by examining the correlation between time-specific AS and MWIS values of the various models in Figure
\ref{fig:metrics-correlation}. We find, for example, no clear association between the consistent
and high MWIS rank of the \texttt{JHUAPL-SLPHospEns} model and its highly variable AS rank. This contrasts 
with a clearly positive correlation between MWIS and AS ranks in other models such as 
\texttt{Karlen-pypm} and \texttt{USC-SI\_kJalpha}.
We also can see in more detail the consistently high performance \texttt{COVIDhub-ensemble} on both scores.
And while it did not submit forecasts for enough weeks to be included in Table \ref{tab:multi-week-performance-summary}, \texttt{CMU-TimeSeries} is an interesting example of a model that 
performed consistently well on AS but had only middling MWIS ranks.

@@

AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of
hospitalizations on January 3, 2021 as well as for the \texttt{per-capita} allocation (Figure \ref{fig:multi-k}A). These
calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of
\Sexpr{format(tot_y, big.mark = ",")} that day. Model ranks by AS were fairly stable across all $K$, and there appears
to be an substantial interval around the observed value over which ranks are constant.

@@

In practice, epidemiological forecasts are often directed at a diverse set of end-users. It may be easy for some of
these users to summarize the consequences of how they use a forecast with a numerical loss function, but for others,
such as officials deciding how best to update public ``situational awareness'' of an outbreak, this may be essentially
impossible. Even those forecast uses that are easily framed as expected loss minimization may differ enough that no
single score would be appropriate for all users. Ideally, targeted forecasting tools could be developed through close
collaboration between modelers and public health officials. However, this may only be possible in settings with
sufficient staffing on both an analytics and a public health team. Increasingly, collaborative modeling hubs are being
used to generate ``one-size-fits-all'' forecasts for many locations at once. But it could still be beneficial in these
settings to expand the set of scoring methods used by the hub to include scores, such as the AS, which are based on
specific public health decision problems.  This could help end users better understand the value of available forecasts
as inputs to their particular decision making contexts. (We note that issues of propriety are raised when forecasts are
evaluated with scoring rules for which they were not elicited or via parameters imputed by a hub after submission, both
of which are done in our application example.  We consider these issues briefly in appendices
\ref{sec:distfromq} and \ref{sec:propriety_of_parametric_approximation}.)

@@

\appendix
\appendixpage
\addappheadtotoc

\section{Introduction}
\label{sec:intro}

We address some technical and methodological points from the main text. We begin in section \ref{sec:proper} by defining proper scoring rules and showing that the allocation score is proper. In section \ref{sec:ex-shortage}, we formalize the concept of a \emph{shortage} of resources, give some key results about expected resource shortages under a distribution characterizing uncertainty about (future) levels of resource need, and illustrate how quantiles arise as the Bayes act in a decision making problem about the quantity of a resource to purchase. Section \ref{sec:bayes-quantiles} gives a justification for the result that the Bayes act for the allocation problem is given by a vector of quantiles of the forecast distributions in each location at a shared probability level, which was stated in section \ref{sec:methods.detailed.specific_allocation} of the main text. We describe the algorithm that we use to compute allocations given a forecast distribution in each location in section \ref{sec:numeric}. In section \ref{sec:distfromq}, we describe the methods for approximating a full distribution from a set of quantiles, implemented in the R package \verb`distfromq`, that we used to support computation of allocation scores from quantile forecasts that were submitted to the US COVID-19 Forecast Hub for the application in the main text. Finally, in section \ref{sec:propriety_of_parametric_approximation} we examine implications for propriety of an analysis that uses summaries of forecasts (such as predictive quantiles) rather than the full forecast distributions for computation of allocation scores.

\section{Proper scoring rules}
\label{sec:proper}

In decision theory, a loss function $l$ is used to formalize a decision problem by assigning numerical value $l(x,y)$ to the
\emph{result} of taking an \emph{action} $x$ in preparation for an \emph{outcome} $y$. A \emph{scoring rule} $S$ is a
loss function for a decision problem where the action is a probabilistic forecast $F$ of the outcome $y$ (or the statement of $F$ by a forecaster).
% As does any loss fun $S$ transforms a random outcome variable $Y$ into a random loss $S(F,Y)$.
We refer to the realized loss $S(F,y)$ as the \emph{score} of $F$ at $y$.

Probabilistic forecasts can be seen as a unique kind of action in that they can be used to generate their
own (simulated) outcome data, against which they can be scored using $S$. A probabilistic forecast
$F$ is thus committed to the ``self-assessment'' $\Ex_F [S(F, Y)] := \Ex [S(F, Y^F)]$, where $Y^F \sim F$ is the random variable defined
by sampling from $F$, as well to an assessment $\Ex_F [S(G, Y)]$ of any alternative forecast $G$.

A natural consistency criterion for $S$ is that, for observations assumed to be drawn from $F$, it will not assess any other forecast $G$
as being better than $F$ itself, that is, that
\begin{align}
\Ex_F [S(F, Y)] \leq \Ex_F [S(G, Y)] \label{eqn:prop_ineq}
\end{align}
for any $F,G$. A scoring rule meeting this criterion is called \emph{proper}. If $S$ were improper, then from the perspective of
a forecaster focussed (solely) on expected loss minimization, the decision to state
a forecast $G$ other than the forecast $F$ which they believe describes $Y$ could be superior to the decision to state $F$.
$S$ is \emph{strictly proper} when
\eqref{eqn:prop_ineq} is a strict inequality, in which case the
\emph{only} optimal decision for a forecaster seeking to minimize their expected loss is to state the forecast they believe to be true.

\subsection{The allocation score is proper}
\label{sec:alloscore_proper}

Our primary decision theoretical procedure, outlined in section \ref{sec:methods.detailed.decisiontheory} of the main text,
uses a decision problem with loss function $s(x,y)$ to define a scoring rule
\begin{align}
S(F,y) := s(x^F,y) \label{eqn:bayes_sr}
\end{align}
where $x^F := \argmin_{x} \Ex_F[s(x,Y)]$ is the Bayes act for $F$ with respect to $s$.
Such scoring rules, which we call \emph{Bayes scoring rules},
are proper by construction since
\begin{align}
\Ex_F [S(F, Y)] &= \Ex_F [ s(x^F, Y) ] \nonumber \\
 &= \mathrm{min}_{x} \Ex_F [ s(x, Y) ] \quad \text{ (by definition of $x^F$)} \\
 &\leq \Ex_F [ s(x^G, Y) ] \label{eqn:dt_proper_key} \\
 &= \Ex_F [ S(G, Y)]. \nonumber
\end{align}

The allocation scoring rule is Bayes and therefore proper.

We note that in the probabilistic forecasting literature (see e.g., \cite{gneiting2011making}, Theorem 3) what we have
termed Bayes scoring rules typically appear via \eqref{eqn:bayes_sr} where $x^F$ is some given functional of $F$ which
can be shown to be \emph{elicitable}, that is, to be the Bayes act for some loss function $s$.
Such a loss function is said to be a \emph{consistent loss (or scoring) function} for the functional $F \mapsto x^F$, and many important
recent results in the literature (e.g., \cite{fisslerziegel2016consistency}) address whether there \emph{exists} any loss
function that is consistent for $x^F$. Our orientation
is different from this insofar as we \emph{begin} by specifying a decision problem and a loss function of subject matter relevance
and use the Bayes act only as a bridge to a proper scoring rule.  Consistency is never in doubt.

\section{Expected shortages}
\label{sec:ex-shortage}

A key feature of loss functions for decision problems used to define quantiles and related scoring rules such as the
CRPS and the WIS (see e.g., \cite{gneiting2011quantiles}, \cite{jose2009evaluating},
and \cite{royset2022optimization}, sections 1.C and 3.C), as well as the allocation loss function presented in this work,
is the presence of a \emph{shortage}:
the amount $\max\{0,y-x\}$ by which a resource demand
$y$ exceeds a supply decision variable $x$, which, for convenience, we write as $(y-x)_{+}$. In particular, a quantile at
probability level $\alpha$ of a distribution $F$ on $\mathbb{R}^1$ (which we assume to have a well-defined density $f(x)$)
is a Bayes act for the loss function
\[
l(x,y) = Cx + L(y-x)_{+}
\]
where $\alpha = 1-C/L$ and $C$ and $L$ can be interpreted as the cost per unit of a resource (such as medicine) and the loss
incurred when a unit of demand (such as illness) cannot be met due to the shortage $(y-x)_{+}$.  This follows because a
Bayes act, as a minimizer of $\Ex_F[l(x,Y)]$, must also be a vanishing point of the derivative
\begin{align}
\dby{}{x} \Ex_F\left[l(x,Y)\right] &= \Ex_F\left[\dby{}{x}l(x,Y)\right] \nonumber\\
&= C + L\Ex_F\left[\dby{}{x}(Y-x)_+\right] \nonumber\\
&= C - L\Ex_F\left[\mathbf{1}\{Y > x\}\right] \nonumber\\
&= C + L(F(x) - 1), \label{eqn:q_deriv}
\end{align}
so that $1-C/L = F(x)$.
The formula $\dby{}{x}\Ex_F\left[(Y-x)_+\right] = F(x) - 1$ for the derivative of the shortage,
used above in \eqref{eqn:q_deriv}, can be obtained from an application of the ``Leibniz Rule'':
\begin{align}
	\frac{d}{dx} \Ex_F [(Y-x)_{+}] &= \frac{d}{dx} \int_{x}^{\infty} (y-x) f_Y(y)dy \nonumber\\
	&= \int_{x}^{\infty} \frac{d}{dx}(y-x) f_Y(y)dy - (x-x) f_Y(x) = -\int_{x}^{\infty} f_Y(y)dy = F(x)-1. \label{eqn:shortage_deriv}
\end{align}
Note that more care is required when $F$ does not have a density.
We will also use this result below in deriving the Bayes act for the allocation loss.

\section{Allocation Bayes acts as vectors of marginal quantiles.}
\label{sec:bayes-quantiles}
